---
title: "lab6"
format: github_document
editor: visual
---

## **Exercise 1: Whale observation (40 min)**

```{r}
# Load required packages
library(tidyverse)
library(knitr)

# Read in the data
whales <- read_csv("https://raw.githubusercontent.com/nt246/NTRES-6100-data-science/main/datasets/whales.csv")
whales |> 
  head() |> 
  kable()
```

#### **Question 1. Create a new data frame that has one row per observer, per species and one single variable of all the information collected. Name this data frame `whales_long`.**

```{r}
whales_long <- whales |>
  pivot_longer(-1, names_to = "Species", values_to = "Observer")
View(whales_long)

```

#### **Question 2. Starting from `whales_long`, create another data frame that includes only events for which there is information. Name this data frame `whales_clean`.**

```{r}
whales_clean <- whales_long |>
  filter(!is.na(Observer))
View(whales_clean)
```

#### **Question 3. Starting from `whales_clean`, create another data frame with one variable per type of information, one piece of information per cell. Some cells might be empty. Name this data frame `whales_split`.**

```{r}
whales_split <- whales_clean |>
  separate(Observer, c("observer", "species", "date", "outcome", "size", "ocean"))
head(whales_split)

```

#### **Question 4. Starting from `whales_split`, create another data frame in which all columns are parsed as instructed below. Name this data frame `whales_parsed`.**

```{r}


```

#### **Question 5. Using `whales_parsed`, print a summary table with: 1) number ship strikes by species, 2) average whale size by species, omitting NA values in the calculation.**

```{r}

```

#### **Question 6. Try to summarize as much information contained in `whales_parsed` as possible in one plot.**

```{r}

```

## **Exercise 2: Baby names (50 min)**

```{r}
# Load required packages
library(babynames) #install.packages("babynames")

babynames |> 
  head() |> 
  kable()

View(babynames)
```

#### **Question 1: What are the 6 most popular boy names and girl names of all time? How has the popularity of each of these names changed over time?**

```{r}
babynames |> 
  filter(sex == "F") |>
  group_by(name) |>
  summarise(total_count=sum(n))
```

#### **Question 2. What are the most popular boy names and girl names in each decade?**

```{r}

```

#### **Question 3. Continue to explore the `babynames` dataset.**

```{r}

```
